### 2014-06-23 dicook/map.Rmd

```{r load_libraries, message=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggvis)
library(maps)
library(ggmap)
library(rworldmap)
library (grid)  	
library (scales)	
```

@@

```{r pairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
library(YaleToolkit)
gpairs(student2012.sub.summary[, c(2, 10, 11)])
# library(cranvas)
# qstudent <- qdata(student2012.sub.summary)
# qscatter(math, read, qstudent)
# qscatter(math, science, qstudent)
# qscatter(read, science, qstudent)
# record_selector(name, qstudent)
# Should do a PCA, to look at what countries do better on what types of math skills
student2012.sub.summary.nomiss <- subset(student2012.sub.summary, !is.na(student2012.sub.summary$Change)&!is.na(student2012.sub.summary$Employ))
rownames(student2012.sub.summary.nomiss) <- student2012.sub.summary.nomiss[,1]
student2012.sub.math.pca <- prcomp(student2012.sub.summary.nomiss[,2:9], scale=T, retx=T)
student2012.sub.math.pca
qplot(PC2, PC3, data=data.frame(student2012.sub.math.pca$x)) + theme(aspect.ratio=1)
# High values on PC2 correspond to high data and interpretation, low corresponds to high spatial 
# Ireland, UK, Greece, US states, Netherland, Norway, NZ, Finland, France, Croatia, Australia come up as high 
# on data and interp but less on spatial AND
# Chine, Albania, Taiwan, Kazakhstan, Korea, Japan, Russia come up as being high on spatial, less on data and interp
rownames(student2012.sub.math.pca$x)[order(student2012.sub.math.pca$x[,2], decreasing=T)]
# High on PC3 correspond to high quantification, employ, low on data
# Israel, Croatia, Estonia, Czech are high; Taiwan, Japan, Indonesia, Massachussetts, Jordan are low
rownames(student2012.sub.math.pca$x)[order(student2012.sub.math.pca$x[,3], decreasing=T)]
student2012.sub.summary.pca <- prcomp(student2012.sub.summary.nomiss[,c(2,10,11)], scale=T, retx=T)
student2012.sub.summary.pca
```

>>>>

```{r load_libraries, message=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(stringr)
library(ggplot2)
library(dplyr)
library(lubridate)
#library(ggvis)
library(maps)
library(ggmap)
library(rworldmap)
library (grid)  	
library (scales)	
```

@@

```{r pairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
library(YaleToolkit)
student2012.sub <- student2012[, c(1:7, seq(501, 550, 5))]
colnames(student2012.sub)[1] <- "name"
student2012.sub$name <- as.character(student2012.sub$name)
# Check mismatches of names
# unique(anti_join(student2012.sub, world.polys)[1])
student2012.sub$name[student2012.sub$name=="Serbia"] <- "Republic of Serbia"
student2012.sub$name[student2012.sub$name=="Korea"] <- "South Korea"
student2012.sub$name[student2012.sub$name=="Chinese Taipei"] <- "Taiwan"
student2012.sub$name[student2012.sub$name=="Slovak Republic"] <- "Slovakia"
student2012.sub$name[student2012.sub$name=="Russian Federation"] <- "Russia"
student2012.sub$name[student2012.sub$name=="Perm(Russian Federation)"] <- "Russia"
student2012.sub$name[student2012.sub$name=="Hong Kong-China"] <- "Hong Kong S.A.R."
student2012.sub$name[student2012.sub$name=="China-Shanghai"] <- "China"
student2012.sub$name[student2012.sub$name=="China-Macau"] <- "China"
student2012.sub$name[student2012.sub$name=="Connecticut (USA)"] <- "United States of America"
student2012.sub$name[student2012.sub$name=="Florida (USA)"] <- "United States of America"
student2012.sub$name[student2012.sub$name=="Massachusetts (USA)"] <- "United States of America"

# Only need one of the plausible values, checked they are effectively identical
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub$PV1MACC <- as.numeric(student2012.sub$PV1MACC)
student2012.sub$PV1MACQ <- as.numeric(student2012.sub$PV1MACQ)
student2012.sub$PV1MACS <- as.numeric(student2012.sub$PV1MACS)
student2012.sub$PV1MACU <- as.numeric(student2012.sub$PV1MACU)
student2012.sub$PV1MAPE <- as.numeric(student2012.sub$PV1MAPE)
student2012.sub$PV1MAPF <- as.numeric(student2012.sub$PV1MAPF)
student2012.sub$PV1MAPI <- as.numeric(student2012.sub$PV1MAPI)
student2012.sub$PV1READ <- as.numeric(student2012.sub$PV1READ)
student2012.sub$PV1SCIE <- as.numeric(student2012.sub$PV1SCIE)
student2012.sub.summary <- summarise(group_by(student2012.sub[,c(1,8:17)], name), math=mean(PV1MATH, na.rm=T), 
                                  mCC=mean(PV1MACC, na.rm=T), mCQ=mean(PV1MACQ, na.rm=T), 
                                  mCS=mean(PV1MACS, na.rm=T), mCU=mean(PV1MACU, na.rm=T), 
                                  mPE=mean(PV1MAPE, na.rm=T), mPF=mean(PV1MAPF, na.rm=T), 
                                  mPI=mean(PV1MAPI, na.rm=T), read=mean(PV1READ, na.rm=T), 
                                  science=mean(PV1SCIE, na.rm=T), 
                                     mathr=diff(range(PV1MATH, na.rm=T)), 
                                  mCCr=diff(range(PV1MACC, na.rm=T)), mCQr=diff(range(PV1MACQ, na.rm=T)), 
                                  mCSr=diff(range(PV1MACS, na.rm=T)), mCUr=diff(range(PV1MACU, na.rm=T)), 
                                  mPEr=diff(range(PV1MAPE, na.rm=T)), mPFr=diff(range(PV1MAPF, na.rm=T)), 
                                  mPIr=diff(range(PV1MAPI, na.rm=T)), readr=diff(range(PV1READ, na.rm=T)), 
                                  sciencer=diff(range(PV1SCIE, na.rm=T)))
colnames(student2012.sub.summary)[3:9] <- c("Change", "Quantity", "Spatial", "Data", "Employ", "Formulate", "Interpret")
gpairs(student2012.sub.summary[, c(2, 10, 11)])
# library(cranvas)
# qstudent <- qdata(student2012.sub.summary)
# qscatter(math, read, qstudent)
# qscatter(math, science, qstudent)
# qscatter(read, science, qstudent)
# record_selector(name, qstudent)
# Should do a PCA, to look at what countries do better on what types of math skills
student2012.sub.summary.nomiss <- subset(student2012.sub.summary, !is.na(student2012.sub.summary$Change)&!is.na(student2012.sub.summary$Employ))
rownames(student2012.sub.summary.nomiss) <- student2012.sub.summary.nomiss[,1]
student2012.sub.math.pca <- prcomp(student2012.sub.summary.nomiss[,2:9], scale=T, retx=T)
student2012.sub.math.pca
qplot(PC2, PC3, data=data.frame(student2012.sub.math.pca$x)) + theme(aspect.ratio=1)
# High values on PC2 correspond to high data and interpretation, low corresponds to high spatial 
# Ireland, UK, Greece, US states, Netherland, Norway, NZ, Finland, France, Croatia, Australia come up as high 
# on data and interp but less on spatial AND
# Chine, Albania, Taiwan, Kazakhstan, Korea, Japan, Russia come up as being high on spatial, less on data and interp
rownames(student2012.sub.math.pca$x)[order(student2012.sub.math.pca$x[,2], decreasing=T)]
# High on PC3 correspond to high quantification, employ, low on data
# Israel, Croatia, Estonia, Czech are high; Taiwan, Japan, Indonesia, Massachussetts, Jordan are low
rownames(student2012.sub.math.pca$x)[order(student2012.sub.math.pca$x[,3], decreasing=T)]
student2012.sub.summary.pca <- prcomp(student2012.sub.summary.nomiss[,c(2,10,11)], scale=T, retx=T)
student2012.sub.summary.pca
```
